---
title: "DATA151: Bivariate Relationships"
author: "STUDENT NAME HERE"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Learning Objectives: 

Students will learn to create scatterplots and describe the relationships between two numeric variables. 

### Example: Take A Hike!

#### STEP 1: Load the data: 

```{r}
body_wgt<-c(120, 187, 109, 103, 131, 165, 158, 116)
backpack_wgt<-c(26, 30, 26, 24, 29, 35, 31, 28)

backpack_df<-data.frame(body_wgt, backpack_wgt)
```

Which variable should be the response and which should be the explantory? 

```{}
## YOUR ANSWER HERE ##
```

#### STEP 2: Scatterplot

```{r warning=FALSE, message=FALSE}
library(tidyverse)

## YOUR CODE HERE ####

```

#### STEP 3: Describe

When looking at a scatterplot you want to be able to describe the overall pattern and for striking departures from that pattern.

You can describe the overall pattern of a scatterplot by the:

* direction – positive or negative
* form – linear or non-linear
* strength – strong (points close together) or weak (points spread out)
* outliers - an individual value that falls outside the overall pattern of the relationship

How would you describe the above scatterplot? 
```{}
### YOUR ANSWER HERE ###
```

#### STEP 4: Correlation Coefficient

```{r}
### YOUR CODE HERE ###
```

What happens when you switch the order of the variables?

```{r}
### YOUR CODE HERE ###
```

#### STEP 5: Activity

**First**: Load in the data

```{r}
data("anscombe")
str(anscombe)
```

**Directions**: 

If your birthday is: 

* January - March: Use variables `x1` and `y1`
* April - June: Use variables `x2` and `y2`
* July - September: Use variables `x3` and `y3`
* October - December: Use variables `x4` and `y4`

Complete the following tasks: 

* Create a scatterplot and describe it
* Calculate the mean and standard deviation for both your `x` and `y` variables
* Calculate the correlation coefficient
* Compare the information you have obtained with your neighbor

```{r}
## SPACE FOR YOUR WORK ##


```


#### STEP 6: Line of Best Fit

```{r}
## YOUR CODE HERE ##
```

What is this line? 

```{r}
# Y~X
## YOUR CODE HERE ##
```











