---
title: 'STAT 441: Random Variables and Distributions in R'
output:
  html_document:
    toc: true
    collapsed: false
    toc_float: true
    toc_depth: 5
  pdf_document:
    toc: true
    toc_depth: '5'
---

## Learning Objectives: 

Students will learn how to ...

* Generate random variables in R from the normal and multivariate normal distributions 
* Calculate probabilities 
* Visualize densities


### Example 1: Normal Distribution

R has several built-in functions for known distributions.  Let’s learn a little bit more about them!

For this example please use the `rnorm()` function.  To learn about the arguments that this function takes you can type `?rnorm` to read the help file. 

#### Step 1: Generate

Generate 10000 observations from the Normal distribution using the defaults in the rnorm function. 

```{r}
### GENERATE

## INSERT CODE HERE ##
```

#### Step 2: Visualize


Create a plot that estimates the density based on these data

```{r warning=FALSE, message=FALSE}
### PLOT DENSITY

## BASE

## INSERT CODE HERE ##

## GGPLOT
library(tidyverse)


## INSERT CODE HERE ##
```


#### Step 3: Density

We can find the height

```{r}
### DENSITY AT 0

## INSERT CODE HERE ##
```

#### Step 4: Probability 

We can evaluate the area under the curve (to the left) 

```{r}
### PROBABILITY

## INSERT CODE HERE ##
```

#### Question: 

What observations do you have about the normal distribution based on your answers above? 


### Example 2: Multivariate Normal Distribution

#### Step 1A: Generate Correlated Data

Generate 10000 observations from the Normal distribution using the defaults in the rnorm function. 

```{r}
### MULTIVARIATE
#install.packages("MASS")
library(MASS)

## GENERATE

## INSERT CODE HERE ##
```


#### Step 1B: Kernel Density

```{r}
### TWO-DIM KERNEL DENSITY
### n=50 GRID


## INSERT CODE HERE ##
```

#### Step 2: Visualize Correlation

Plots built into base R: 

```{r}
### PLOT (SCATTER)

## INSERT CODE HERE ##

### CONTOUR

## INSERT CODE HERE ##

### HEATMAP

## INSERT CODE HERE ##

### PERSPECTIVE

## INSERT CODE HERE ##
```

#### Step 3A: Generate Independent Data

Generate data from the multivariate normal distribution that has zero correlation 

```{r}
## ZERO CORR 

## INSERT CODE HERE ##
```


#### Step 3B: Kernel Density

```{r}
### ESTIMATED DENSITY

## INSERT CODE HERE ##
```

#### Step 4: Visualize Independence

Plots built into base R: 

```{r}
### PLOT (SCATTER)

## INSERT CODE HERE ##

### CONTOUR

## INSERT CODE HERE ##

### HEATMAP

## INSERT CODE HERE ##

### PERSPECTIVE

## INSERT CODE HERE ##
```

#### Question: 

Compare the sets of plots.  What do you observe? 