---
title: 'STAT 441: Linear Algebra in R'
output:
  pdf_document:
    toc: true
    toc_depth: '5'
  html_document:
    toc: true
    collapsed: false
    toc_float: true
    toc_depth: 5
---

### Learning Objectives 

Students will learn how to ...

* Construct diagnostic plots to explore model assumptions using residuals
  



### A. Import the Data 

These data are from Open Intro.  

Documentation in R: 

Zagat is a public survey where anyone can provide scores to a restaurant. The scores from the general public are then gathered to produce ratings. This dataset contains a list of 168 NYC restaurants and their Zagat Ratings.

```{r}
### IMPORT THE DATA
### NYC RESTAURANT RATINGS
#install.packages("openintro")
library(openintro)

data("nyc")

head(nyc)
str(nyc)
#dim(nyc)
```

### B. Visualize 

Create a pairs plot to compare all two way relationships. 

```{r}
### PAIRS PLOT
library(tidyverse)

## INSERT CODE HERE ##
```


### C. Response Vector

Create a response vector. 

```{r}
### RESPONSE VECTOR
## INSERT CODE HERE ##
```

### D. Design Matrix

The design matrix always has a column of one's to start, followed by the explanatory variables. 

```{r}
### DESIGN MATRIX
## INSERT CODE HERE ##
```

### E. Transpose

The transpose of a matrix swaps the rows and columns. 

```{r}
### TRANSPOSE
## INSERT CODE HERE ##
```


### F. Matrix Multiplication 

We use the `%*%` to get the dot product of matrices. 

```{r}
### MATRIX MULTIPLICATION
## INSERT CODE HERE ##
```

### G. Matrix Inverse

The inverse of a matrix is given by `solve`. 

Note: We need a square matrix. 

```{r}
### INVERT
## INSERT CODE HERE ##
```

Check our work: 

```{r}
### CHECK FOR IDENTITY
## INSERT CODE HERE ##
```

### H. Estimate Vector

```{r}
### BRING C THROUGH E TOGETHER
## INSERT CODE HERE ##
```

### I. CHECK

```{r}
### FIT MOD
## INSERT CODE HERE ##
```

