---
title: 'STAT 441: Introduction to Linear Models'
output:
  html_document:
    toc: true
    collapsed: false
    toc_float: true
    toc_depth: 5
  pdf_document:
    toc: true
    toc_depth: '5'
---

## Learning Objectives: 

In this lesson students will learn ... 

* The basics of R
  * Random variable functions built into R
  * Variable assignment 
  * How to use simple mathematical functions in R to define relationships
* Plotting in base R
* Optional: plotting with ggplot

## Example 1: Creating Linear Relationships

### 1. Variable Assignment 

In R you can use the equals sign `=` or the arrow `<-` to do variable assignment. 


### 2. Generating Data in R 

There are several distributions built into base R, from which you can draw values. 

#### Uniform Distribution

Learn about the `runif` function

```{r}
### HELP
## INSERT CODE HERE ##
```

##### Question 

What are the arguments for the `runif` function?  What are the default values?

```{}
## YOUR NOTES HERE ##
```

##### Activity

###### Step 1

Generate 100 random numbers from the interval (0,1) and call it `X`

```{r}
### Generate 100 random numbers from the interval (0,1) and call it X

## INSERT CODE HERE ##
```

Create a histogram to look at the data. 

```{r}
### Histogram in Base R
## INSERT CODE HERE ##
```

###### Step 2

Generate 100 random numbers from the interval (0,1) and call it W

```{r}
### Generate 100 random numbers from the interval (0,1) and call it W

## INSERT CODE HERE ##
```


###### Step 3

Define Y to be a linear combination of X and W, where 
$$Y=0.15X + 0.85W$$

```{r}
### Define Y to be a linear combination of X and W

## INSERT CODE HERE ##
```

###### Step 4

Make a scatterplot of (X,W)

```{r}
### SCATTERPLOT IN BASE R
## INSERT CODE HERE ##
```

##### Question

How would you describe the relationship between X and W?

```{}
## YOUR NOTES HERE
```

###### Step 5

Make a scatterplots of (X, Y) and (W, Y)

```{r}
### SCATTERPLOTS OF X,Y and W,Y
## INSERT CODE HERE ##
```

##### Question

Are these plots similar or different? Do you have a hypothesis about why
they might look different?

```{}
## YOUR NOTES HERE
```


## Example 2: Linear Models

Consider the following examples of linear models: 

* Simple Linear Regression: $Y=\beta_0+\beta_1 X$
* Polynomial Regression: $Y=\beta_0+\beta_1 X+ \beta_2 X^2$
* Log Transformation: $Y=\beta_0+\beta_1 ln(X)$
* Multiple Linear Regression: $Y=\beta_0+\beta_1 X_1+\beta_2 X_2$

##### Activity

Program examples of the models above and in R. Then create a scatter plot to visualize the
relationship. Compare and contrast plots.

```{r}
### SIMPLE LINEAR
## INSERT CODE HERE ##

### PARAMETERS
## INSERT CODE HERE ##

### MODEL
## INSERT CODE HERE ##

### PLOT
## INSERT CODE HERE ##
```

```{r}
### POLYNOMIAL LINEAR
## INSERT CODE HERE ##

### PARAMETERS
## INSERT CODE HERE ##

### MODEL
## INSERT CODE HERE ##

### PLOT
## INSERT CODE HERE ##
```

```{r}
### LOG TRANSFORMED
## INSERT CODE HERE ##

### PARAMETERS
## INSERT CODE HERE ##

### MODEL
## INSERT CODE HERE ##

### PLOT
## INSERT CODE HERE ##
```

```{r}
### MULTIPLE
## INSERT CODE HERE ##

### PARAMETERS
## INSERT CODE HERE ##

### MODEL
## INSERT CODE HERE ##

### PLOT
## INSERT CODE HERE ##
```
