---
title: 'STAT 441: Application of SLR Estimates'
output:
  html_document:
    toc: true
    collapsed: false
    toc_float: true
    toc_depth: 5
  pdf_document:
    toc: true
    toc_depth: '5'
---

## Learning Objectives: 

Students will learn how to ...

* Import data into R 
* Obtain summary statistics in R 
* Create scatter plots
* Perform mathematical functions on vectors of data in R (elementwise)


### Example 1: Ice Cream 

Last week was hot!!! What refreshing treat do people enjoy on a hot day?  ICE CREAM! 
As we come to the end of the summer and the cooler weather of the fall, let’s take a look at this fun dataset about ice cream sales and weather. 

These data are provided on github by Guilherme D. F. Silva (guilhermedom)
The following directions are in base R, but can be done within the tidyverse as well. 

#### Step 1: Import the Data

```{}
###Import the data
icecream<-read.csv("PATHWAY HERE",
                 header=TRUE)
```

#### Step 2: Look at the Structure of the Data

```{r}
###Look at the data

## INSERT YOUR CODE HERE ###
```

#### Step 3: Save Vectors for Response and Explanatory

```{r}
###Get access to the data set and create variables for easy access later
## INSERT YOUR CODE HERE ###
```

#### Step 4: Create a Histogram

```{r}
### IN BASE
###Make a histogram of the response variable 
## INSERT YOUR CODE HERE ###


### IN GGPLOT
library(tidyverse)

## INSERT YOUR CODE HERE ###
```


#### Step 5: Summary Statistics

```{r}
###Report basic statistics of the response variable 
## INSERT YOUR CODE HERE ###
```

#### Step 6: Linear Model in R

```{r}
###Run a simple linear regression model on shear against age and store the results
## INSERT YOUR CODE HERE ###
```


#### Step 7: Scatter Plot

```{r}
####Create a scatterplot
## BASE
## INSERT YOUR CODE HERE ###
####Add the least squares regression line
## INSERT YOUR CODE HERE ###

## GGPLOT
## INSERT YOUR CODE HERE ###
```

#### Step 8: Verify from Scratch 

Use the mathematical equations derived in class to confirm the least squares estimates from lm.

```{r}
## INSERT YOUR CODE HERE ###
```


### Example 2: Rockets

The rocket data are available here: 

```{r}
rocket<-read.csv("https://raw.githubusercontent.com/kitadasmalley/Teaching_Public/refs/heads/main/STAT441/FA25/data/rocketProp.csv",
                 header=TRUE)
```




