---
title: 'STAT 441: Midterm 1'
author: "NAME HERE"
date: "DATE"
output: pdf_document
---

## Application in R

For this part of the exam you will need to use the fruitohms dataset in the DAAG package.  A data frame with 128 observations on apparent juice content (percent) and electrical resistance (in ohms). 

### Description in R: Electrical Resistance of Kiwi Fruit 

Data are from a study that examined how the electrical resistance of a slab of kiwifruit changed with the apparent juice content.

### Access the Data

```{r}
# Import the package
install.packages("DAAG")
library(DAAG)

# Load in the data
data("fruitohms")

# Learn about the data
?fruitohms

```

### Part A: Variable (5 points)

Identity the explanatory variable and the response variable for this study. 


### Part B: Visualize (5 points)

Use R to create a scatter plot of the two variables.  Examine the scatter plot and verbally describe the overall relationship (linear or nonlinear, positive or negative).

```{r}
### INSERT CODE HERE ###
```

### Part C: Model (5 points)

Create a simple linear model and add the least squares regression line to the scatter plot.  What is the equation of the line? 

### Part D: Interpret (2 points)

Interpret the slope in the context of these data. 

### Part E: Hypothesis Testing (5 points)

Perform a hypothesis test for the slope.  Provide a five part conclusion for the hypothesis test for slope using the summary model output. 

```{r}
### MODEL SUMMARY

## INSERT CODE HERE ##
```


### Part F: Critical Thinking (3 points)

Would you trust this model, given your observations in Part B?

