---
title: 'DATA 151: Project Milestone #6'
author: "INSERT GROUP MEMBERS HERE"
date: "DATE HERE"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Relationships Between Two Numeric Variables


### STEP 0: Import the data


```{r}
## CALL THE DATA IN 

## INSERT CODE HERE ##
```


### Motivating Question 

State a question of interest for relationships between two numeric variables.
Identify a numeric response variable in your dataset and a numeric explanatory variable.

STATE YOUR QUESTION OF INTEREST HERE. 

### STEP 1: Create a scatter plot.

For full points, make sure to label the axes and give the plot a title

```{r}
## MAKE A SCATTERPLOT

## INSERT CODE HERE ##
```

### STEP 2: Describe the relationship between two numeric variables

  - Direction
  - Form 
  - Strength 
  - Outliers 

### STEP 3: Add a smooth to the scatter plot 

Using method="lm"

```{r}
## ADD THE SMOOTH TO YOUR SCATTPLOT

## INSERT CODE HERE ##
```

### STEP 4: Fit a line of best fit to your data using the lm function. 

Write the equation for the line.

```{r}
## LM
## INSERT CODE HERE ##

## TYPE EQUATION HERE
```


### Bonus points: Include a categorical variable to color your plot and look for subgroupings

Should have two numeric variables (x and y) and one categorical variable for color

```{r}
## BONUS

## INSERT CODE HERE ##
```


