---
title: "DATA 151: Project Milestone #4"
author: "INSERT GROUP MEMBERS HERE"
date: "DATE HERE"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating Tables and Bar Graphs


### STEP 0: Import the data


```{r}
## CALL THE DATA IN 

## INSERT CODE HERE ##
```


### PART 1: Motivating Question 1

State a question of interest for one categorical variable. 

STATE YOUR QUESTION OF INTEREST HERE. 

### STEP 1: One-way Table (One Categorical Variable)

```{r}
## MAKE A ONE-WAY TABLE

## INSERT CODE HERE ##
```

### STEP 2: Bar Graph

```{r}
## MAKE A UNIVARIATE BAR GRAPH

## INSERT CODE HERE ##
```

### STEP 3: Pie Chart

```{r}
## MAKE A PIE CHART

## INSERT CODE HERE ##
```

### STEP 4: Discuss your findings

PLEASE DISCUSS YOUR FINDINGS HERE

### PART 2: Motivating Question 2

State a question of interest exploring the relationship between two categorical variables

STATE YOUR QUESTION OF INTEREST HERE. 

### STEP 1: Two-way Table (Two Categorical Variables)

```{r}
## MAKE A TWO-WAY TABLE

## INSERT CODE HERE ##
```

### STEP 2: Stacked Bar Graph

```{r}
## MAKE A STACKED BAR GRAPH

## INSERT CODE HERE ##
```

### STEP 3: Side-by-side Bar Graph

```{r}
## MAKE A SIDE-BY-SIDE BAR GRAPH

## INSERT CODE HERE ##
```

### STEP 4: Filled Bar Graph

What is the connection to the conditional distribution? 

```{r}
## MAKE A FILLED BAR GRAPH

## INSERT CODE HERE ##
```

### STEP 5: Discuss your findings

PLEASE DISCUSS YOUR FINDINGS HERE